---
title: "<center>AEEIP_vignette</center>"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
abstract: |
  To evaluate antibody quality,we propose the novel Antibody Efficiency Estimation from 
  Immunoprecipitation data (AEEIP) model. It contains a maximum a posterior (MAP) estimate
  the efficiency of antibody. It is able to calculate the proportion of mixed Input
  sample in IP sample. It also try to change the hyper parameter of beat prior to
  get the better result.Please don't hesitate to contact <Haozhe.Wang17@student.xjtlu.edu.cn> 
  if you have any questions.\n <center> **Table of content** </center>
vignette: >
  %\VignetteIndexEntry{AEEIP_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  % \VignetteDepends{ggplot2} 
  % \VignetteDepends{DESeq2} 
  % \VignetteDepends{ggVennDiagram}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# install.packages("devtools")
# library("devtools")
# install_github("whz991026/AEEIP")

library(AEEIP)
```

## load other libraries
Load the other packages:

* library(ggVennDiagram)
* library(DESeq2)
```{r, include=FALSE}


# install.packages("ggVennDiagram")

library(ggVennDiagram)
library(DESeq2)

#    install.packages("ggplot2")

library(ggplot2)
```

## 0.1
### simulate data
This section is getting the simulate data with 0.1 mix rate and replicates is 3
```{r}

# the mix rate is 0.1
set.seed(1)
data <-simulateData(T=10000, hyper_a_1=20,hyper_b_1 =0.2, hyper_a_2=50,hyper_b_2 =0.05,
pure_rate=0.99,mix_rate=0.1,mode=2,replicates=3)

```


### corrected the bias

#### replicates 1
corrected the bias for the replicates 1
```{r}

set.seed(1)
# firstly we set the initial tau is near to 0 and let the size of beta to be large enough
# to see the plot
res <- AEEIP_test(r_t_pure=data[[3]][[1]],r_t_mix=data[[4]][[1]],beta=sum(data[[4]][[1]])*1e1,alpha_a=NA,
alpha_b=NA,tau_a_vector=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
max_iteration =1e6,stop_stepsize=1e-7)

res1 <- AEEIP_test_final(res,r_t_pure=data[[3]][[1]],r_t_mix=data[[4]][[1]],beta=sum(data[[4]][[1]])*1e1,alpha_a=NA,                      alpha_b=NA,tau_a=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
                     max_iteration =1e6,stop_stepsize=1e-7)

res_replicate1 <- res1[[1]]
data.frame_1 <- res1[[2]]
data.frame_2 <- res1[[3]]

plot1 <- ggplot() + geom_point(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_bayes`))+
     geom_vline(aes(xintercept=0.1), colour="#BB0000", linetype="dashed")+
     geom_line(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_no_bayes`), colour="blue")

plot2 <- ggplot() + geom_point(data=data.frame_2,aes(x=tau_a_no_bayes,y=diff))+
     geom_hline(aes(yintercept=0), colour="#BB0000", linetype="dashed")

res_replicate1$tau_a_bayes
```



#### replicates 2
```{r}

set.seed(1)
# firstly we set the initial tau is near to 0 and let the size of beta to be large enough,
# to see the plot
res <- AEEIP_test(r_t_pure=data[[3]][[2]],r_t_mix=data[[4]][[2]],beta=sum(data[[4]][[2]])*1e1,alpha_a=NA,
alpha_b=NA,tau_a_vector=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
max_iteration =1e6,stop_stepsize=1e-7)


res1 <- AEEIP_test_final(res,r_t_pure=data[[3]][[2]],r_t_mix=data[[4]][[2]],beta=sum(data[[4]][[2]])*1e1,alpha_a=NA,                      alpha_b=NA,tau_a=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
                     max_iteration =1e6,stop_stepsize=1e-7)

res_replicate2 <- res1[[1]]
data.frame_1 <- res1[[2]]
data.frame_2 <- res1[[3]]

plot1 <- ggplot() + geom_point(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_bayes`))+
     geom_vline(aes(xintercept=0.1), colour="#BB0000", linetype="dashed")+
     geom_line(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_no_bayes`), colour="blue")

plot2 <- ggplot() + geom_point(data=data.frame_2,aes(x=tau_a_no_bayes,y=diff))+
     geom_hline(aes(yintercept=0), colour="#BB0000", linetype="dashed")

res_replicate2$tau_a_bayes
```




#### replicates 3
```{r}

set.seed(1)
# firstly we set the initial tau is near to 0 and let the size of beta to be large enough,
# to see the plot
res <- AEEIP_test(r_t_pure=data[[3]][[3]],r_t_mix=data[[4]][[3]],beta=sum(data[[4]][[3]])*1e1,alpha_a=NA,
alpha_b=NA,tau_a_vector=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
max_iteration =1e6,stop_stepsize=1e-7)

res1 <- AEEIP_test_final(res,r_t_pure=data[[3]][[3]],r_t_mix=data[[4]][[3]],beta=sum(data[[4]][[3]])*1e1,alpha_a=NA,                      alpha_b=NA,tau_a=0,l=1,prop_vector=seq(0.01,0.99,by=0.01),
                     max_iteration =1e6,stop_stepsize=1e-7)

res_replicate3 <- res1[[1]]
data.frame_1 <- res1[[2]]
data.frame_2 <- res1[[3]]

plot1 <- ggplot() + geom_point(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_bayes`))+
     geom_vline(aes(xintercept=0.1), colour="#BB0000", linetype="dashed")+
     geom_line(data=data.frame_1,aes(x=`tau_a_no_bayes`,y=`tau_a_no_bayes`), colour="blue")

plot2 <- ggplot() + geom_point(data=data.frame_2,aes(x=tau_a_no_bayes,y=diff))+
     geom_hline(aes(yintercept=0), colour="#BB0000", linetype="dashed")

res_replicate3$tau_a_bayes
```

### use the non_mix real data to run the DESeq2
use DESeq2 to analysis the real data, to see the differential between the input and IP.
```{r}
    counts <- cbind(as.data.frame(data[[1]]),as.data.frame(data[[2]]))
    design = data.frame(
                    Trt = c(rep("input", dim(as.data.frame(data[[1]]))[2]),
                            rep("IP", dim(as.data.frame(data[[1]]))[2])))
    model = ~  Trt 
    # 
    dds <- DESeq2::DESeqDataSetFromMatrix(countData = counts,
                                colData = design,
                                design = model)


    data_size <- as.matrix(counts)


    # first log
    log_data <- log(data_size)
    log_data [is.infinite(log_data)] <- NA
    log_mean <- rowMeans(log_data)
    log_s <- log_data-log_mean

    # then exp
    s_size <- exp(apply(log_s,2,function(x)median(x,na.rm=TRUE)))
    names(s_size) <- NULL
    dds$sizeFactor <- s_size
    # DESeq2::sizeFactors(dds) = s_size

    dds <- DESeq2::DESeq(dds, test = "Wald")
    res_DESeq2_real = results(dds)#, name = "Trt_IP_vs_input"
    
    
    res_DESeq2_real_01_1 <- res_DESeq2_real
```

### use the mix data to run the DESeq2
use DESeq2 to analysis the mix data, to see the differential between the input and IP.
```{r}
    counts <- cbind(as.data.frame(data[[3]]),as.data.frame(data[[4]]))
    design = data.frame(
                    Trt = c(rep("input", dim(as.data.frame(data[[3]]))[2]),
                            rep("IP", dim(as.data.frame(data[[3]]))[2])))
    model = ~  Trt 
    # 
    dds <- DESeqDataSetFromMatrix(countData = counts,
                                colData = design,
                                design = model)
 
    
    data_size <- as.matrix(counts)
  
  
    # first log
    log_data <- log(data_size)
    log_data [is.infinite(log_data)] <- NA
    log_mean <- rowMeans(log_data)
    log_s <- log_data-log_mean
    
    # then exp
    s_size <- exp(apply(log_s,2,function(x)median(x,na.rm=TRUE)))
    names(s_size) <- NULL
    sizeFactors(dds) = s_size
    
    dds <- DESeq(dds, test = "Wald")
    res_DESeq2_mix = DESeq2::results(dds)
    
    res_DESeq2_mix_01_1 <- res_DESeq2_mix
```

### use the corrected mix data to run the DESeq2 
```{r}
 
input <- as.data.frame(data[[3]])
IP <- as.data.frame(data[[4]])

res_DESeq2_mix_corrected <- DESeq2_AEEIP(list(res_replicate1,res_replicate2,res_replicate3),input,IP)
```


### evaluate
```{r}


# we see the difference is small for the corrected mix data
 

# to see the i=relation between the three results
index_real <- which(res_DESeq2_real$pvalue<=5e-2)
threshold1 <- res_DESeq2_mix_corrected$pvalue[order(res_DESeq2_mix_corrected$pvalue)[length(index_real)]]
threshold2 <- res_DESeq2_mix$pvalue[order(res_DESeq2_mix$pvalue)[length(index_real)]]
index_mix_corrected <- which(res_DESeq2_mix_corrected$pvalue<=threshold1)
index_mix <- which(res_DESeq2_mix$pvalue<=threshold2)
 
index_list <- list(real=index_real,corrected=index_mix_corrected,
                   mix=index_mix)
 

ggVennDiagram(index_list,category.names = c("real","corrected","mix"),
              label = "count", 
              label_color = "black",
              label_alpha = 0,
              edge_lty = "dashed", 
              edge_size = 1) +
  scale_fill_gradient(low="white",high = "#b9292b",name = "gene count")




# we can see the intersection between real data and corrected mix data is better than the mix data.


```

